Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-22 ASA2013/asa2013.tex
la2pdf commit - 2013:01:22	22:58:06	(Tue22.Jan)

The good news is that this step can greatly benefit from both arithmetic optimizations and parallel computing power. By carefully evaluating (\ref{spatialR}) we can see that:
\begin{itemize}
\item 
\end{itemize}

@@

For larger systems the inversion step will completely dominate the computation time. Although the dimensionality of the coveriance matrix was reduced to $L$ due to subarray averaging, we can reduce it further by moving to the frequency domain. If the data is narrowband enough there is a very close relation between the frequency conent and angles of arrival in the spatial domain; the zero-frequency component will correspond to broadside, and    component corresponding to broadside and increasing frequencies  This is usually handled by reducing the dimensionality of the covariance matrix to

@@

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1. \label{lca_def}
\end{gather}
Note the similarity between (\ref{mvdr_def}) and (\ref{lca_def}). The LCA and MVDR uses essentially the same optimization criterion, but whereas MVDR searches a continuous solution space for the best set of weights to apply, the LCA beamformer only have $P$ solutions to select from. However, once the LCA have 10-20 well designed windows to choose from the its ability to create high quality images quickly approaches the quality level of the MVDR beamformer.

@@

Second we assign $L$ threads per pixel to traverse the diagonals of $\eR$. On the top row a full computation is carried out, then that entire row is saved back to global memory following a collective access pattern that maximize global memory bandwidth. For subsequent rows the threads move along the diagonals while performing iterative summations; the result from the previous element on the diagonal is updated by adding and removing the correlation coefficients that enters and exits the sum, respectively. To minimize memory consumption, we compute the coefficients again every time we need them. When a thread has finished up a diagonal, we have them wrap around to compute one of the diagonals in the lower triangular of $\eR$. Since $\eR$ is conjugate symmetric, the values in the leftmost column is obtained by a complex conjugate copy of the relevant value in the first row. Combined these steps balances the load evenly on all threads, is almost completely free of arithmetic redundancy, and consumes less memory.

>>>>

First, we target this limitation arithmetic optimizations and parallel computing power. In particular, we may infer from (\ref{spatialR}) that:
\begin{itemize}
\item The subarray covariance matrices $R_l[n] = \x_l[n]\x_l\H[n]$ are Hermitian semi-definite, which means that we can get away with computing only one triagonal in them.
\item When summing $R_l[n]$ most of the multiplications can be computed only once, and most sums need only be updated iteratively.
\end{itemize}

@@

As we increase system size the inversion step will become the main bottleneck. Although the dimensionality of the covariance matrix was reduced to $L$ due to subarray averaging, we can reduce it further by moving to the frequency domain. For narrowband  relation between the frequency conent and angles of arrival in the spatial domain; the zero-frequency component will correspond to broadside, and larger frequency components will correspond to increasing .   component corresponding to broadside and increasing frequencies  This is usually handled by reducing the dimensionality of the covariance matrix to

@@

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1. \label{lca_def}
\end{gather}
Note the similarity between (\ref{mvdr_def}) and (\ref{lca_def}). The LCA and MVDR uses essentially the same optimization criterion, but whereas MVDR searches a continuous solution space for the best set of weights to apply, the LCA beamformer only have $P$ solutions to select from. However, we have found that once the LCA has 10-20 well designed windows available the quality of the images it produces quickly approaches that of the MVDR beamformer.

@@

Second we assign $L$ threads per pixel to traverse the diagonals of $\eR$. On the top row a full computation is carried out, then that entire row is saved back to global memory following a collective access pattern that maximizes global memory bandwidth. For subsequent rows the threads move along the diagonals while performing iterative summations; the result from the previous element on the diagonal is updated by adding and removing the correlation coefficients that enters and exits the sum, respectively. To minimize memory consumption, we compute the coefficients again every time we need them. When a thread has finished up a diagonal, we have them wrap around to compute one of the diagonals in the lower triangular of $\eR$. Since $\eR$ is conjugate symmetric, the values in the leftmost column is obtained by a complex conjugate copy of the relevant value in the first row. Combined these steps balances the load evenly on all threads, is almost completely free of arithmetic redundancy, and consumes less memory.
